---
title: "Class10: Comparative Structure Analysis (Pt2)"
author: "Wanning Cui"
format: pdf
---
Today we are going to finish off Lab 10 on learning about and analyzing protein structures.

##Comparative structure analysis of Adenylate Kinase

We will use the `bio3d` package for this analysis that starts with a single sequence.

We will also use the `msa` package from BioConductor.

First we need to install the `BiocManager` we install it the same way as other packages with `install.packages()` command.

We use `BiocManager::install()` to install any other Bioconductor package we want - like `msa` in this case.

```{r}
library(bio3d)
```

```{r}
library(BiocManager)
```

First we will pick a sequence of interest - I will pick "1ake_A"

```{r}
aa <- get.seq("1ake_A")
aa
```

I want to search for all related structures in the PDB database.

```{r}
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

```{r}
head(b$hit.tbl)
```

```{r}
hits$pdb.id
```

>Side-note: let's save our results thus for so we don't have to run blast again.

```{r}
save(hits,b,file="blast_results.Rds")
```

```{r}
load("blast_results.Rds")
hits
```
```{r}
#l eval:false
hits <- plot(b)
```

Now we will download all these related structures from the database with `get.pdb()`

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

##Align and superpose structures

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```
```{r}
pdbs
```
## Principal Component Analysis

We can do a PCA on the coordinate data (i.e. the `$xyz` values) of all these structures.

```{r}
pc.xray <- pca(pdbs)
plot(pc.xray)
```
```{r}
dim(pdbs$xyz)
```

```{r}
plot(pc.xray,1:2)
```
```{r}
mktrj(pc.xray,file="pca_results.pdb")
```


