---
title: "class09"
author: "Wanning Cui"
format: pdf
---

```{r}
fna.data <- "WisconsinCancer.csv"
```

```{r}
wisc.df <- read.csv(fna.data, row.names=1)
head(wisc.df)
```
```{r}
skimr::skim(wisc.df)
```
Store diagnosis column for later use we will exclude this from our dataset for analysis
```{r}
#CREATE DIAGNOSIS VECTOR FOR LATER
diagnosis <- as.factor (wisc.df$diagnosis)
wisc.data <- wisc.df[,-1]
```

>Q1. How many observations are in the dataset?

```{r}
nrow(wisc.df)
```
>Q2. How many of the observations have a malignant diagnosis?

```{r}
sum(wisc.df$diagnosis=="M")
```
```{r}
table(wisc.df$diagnosis)
```


>Q3. How many variables/features in the data are suffixed with _mean?

```{r}
grep("_mean",colnames(wisc.df),value=T)
#colnames(wisc.df)
```

#2. Principal Component Analysis

We need to use `scale=True` here as shown above with our `skim()` report. We could also look at the sd and mean of our columns and see they are on very different scales.

```{r}
# Check column means and standard deviations
colMeans(wisc.data)
apply(wisc.data,2,sd)
```

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp(wisc.data,scale=TRUE)
summary(wisc.pr)
```
>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

```{r}
v <- summary(wisc.pr)
pcvar <- v$importance[3,]
pcvar ["PC1"]
```

0.4427

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

```{r}
which(pcvar >= 0.7)[1]
```
3 PCs

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

```{r}
which(pcvar >= 0.9)[1]
```

7 PCs

>Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

```{r}
biplot(wisc.pr)
```
This is not easy to understand because the relationship is not clear at all and the data is hard to interpret.

```{r}
plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnosis)
```
Most red points (malignant diagnosis) are on one side whereas the black points are clustered more tightly on the other side. There seem to be a more positive correlation between PC1 and PC2.

```{r}
plot( wisc.pr$x, col = diagnosis , 
     xlab = "PC1", ylab = "PC2")
```

>Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,3],col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```

```{r}
# Create a data.frame for ggplot
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
```
```{r}
library(ggplot2)
```
```{r}
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```

#Variance explained

```{r}
# Calculate variance of each component
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

```{r}
# Variance explained by each principal component: pve
pve <- pr.var / sum(pr.var)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```
```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```
```{r}
library(factoextra)
fviz_eig(wisc.pr, addlabels = TRUE)
```

#Communicating PCA results
>Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean? This tells us how much this original feature contributes to the first PC.

```{r}
wisc.pr$rotation[,1]
```
The component is -0.26085376

#3. Hierarchical clustering

```{r}
# Scale the wisc.data data using the "scale()" function
data.scaled <- scale(wisc.data)
```
```{r}
data.dist <-dist(data.scaled)
```
```{r}
wisc.hclust <- hclust(data.dist,method="complete")
```

>Q10. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

```{r}
plot(wisc.hclust)
abline(wisc.hclust, col="red", lty=2)
```
The height is 20

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4, h=NULL)
```
```{r}
table(wisc.hclust.clusters, diagnosis)
```

>Q12. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.

I also prefer the "ward.D2" method because it results in minimal variance from different clusters.

#4. Combining methods

```{r}
wisc.pr.hclust <- hclust (dist(wisc.pr$x[,1:7]),method="ward.D2")
plot(wisc.pr.hclust)
abline(wisc.pr.hclust, col="red", lty=2)
```

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
table(grps)
```
```{r}
table(grps, diagnosis)
```
```{r}
plot(wisc.pr$x[,1:2], col=grps)
```
```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
```
```{r}
g <- as.factor(grps)
levels(g)
```
```{r}
g <- relevel(g,2)
levels(g)
```
```{r}
# Plot using our re-ordered factor 
plot(wisc.pr$x[,1:2], col=g)
```
```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
d.pc <- dist(wisc.pr$x[,1:7])
wisc.pr.hc <- hclust(d.pc, method="ward.D2")
plot(wisc.pr.hc)
```

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hc, k=4)
table(wisc.pr.hclust.clusters, diagnosis)
```

```{r}
table(diagnosis,wisc.pr.hclust.clusters)
```

>Q13. How well does the newly created model with four clusters separate out the two diagnoses?

It does not help with separating the diagnoses because the false positive or non-significant results are not removed. 

>Q14. How well do the hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.

```{r}
wisc.pr.hc <- hclust(d.pc, method="single")
wisc.pr.hclust.clusters <- cutree(wisc.pr.hc, k=4)
table(diagnosis,wisc.pr.hclust.clusters)
```
```{r}
wisc.pr.hc <- hclust(d.pc, method="average")
wisc.pr.hclust.clusters <- cutree(wisc.pr.hc, k=4)
table(diagnosis,wisc.pr.hclust.clusters)
```
There's not much difference among these hierarchical clustering models.


#6. Prediction

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```
```{r}
plot(wisc.pr$x[,1:2], col=g)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

>Q16. Which of these new patients should we prioritize for follow up based on your results?

We should prioritize the patients with malignant diagnoses from cluster 2.

```{r}
sessionInfo()
```

