---
title: "Class 5: Data Visualization with ggplot"
author: "Wanning"
format: pdf
---

##Using GGPLOT

To use ggplot2 we first need to install it on our computers. To do this we will use the function `install.packages()`.

Before I use any package functions I have to load them up with a `library()` call, like so:

```{r}
library(ggplot2)
ggplot(cars)
```
```{r}
head(cars)
```
There is always the "base R" graphics system, i.e. `plot()`

```{r}
plot(cars)
```
To use ggplot I need to spell out at least 3 things:
-data (the stuff I want to plot as a data.frame)
-aesthetics(aes()values - how the data map to the plot).
-geoms (how I want things drawn)

```{r}
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  geom_smooth(method="lm")
```
```{r}
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_line()
```
```{r}
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  geom_smooth(method="lm", se=FALSE)
```
```{r}
ggplot(cars)+
  aes(x=speed, y=dist)+
  geom_point()+
  labs(title="Speed and Stopping Distances of Cars",
       x="Speed(MPH)",
       y="Stopping Distance(ft)",
       subtitle= "Your informative subtitle text here",
       caption="Dataset:'cars'")+
  geom_smooth(method="lm",se=FALSE)+
  theme_bw()
```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
The number of genes in this dataset is:

```{r}
nrow(genes)
```

The column names are:

```{r}
colnames(genes)
```

The number of columns is:

```{r}
ncol(genes)
```

The number of 'up'regulated genes is:
```{r}
table(genes$State)
```
There are `r sum(genes$State == "up")` "UP" regulated genes in this dataset


The fraction of total genes that is up-regulated in this dataset:

```{r}
127/(72+4997+127)
```
```{r}
ggplot(genes)+
  aes(x=Condition1, y=Condition2)+
  geom_point()
```
```{r}
p <- ggplot(genes)+
  aes(x=Condition1, y=Condition2, col=State)+
  geom_point()
```

```{r}
p
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red"))+
  labs(title="Gene Expression Changes Upon Drug Treatment",
       x="Control(no drug)",
       y="Drug Treatment")
```
```{r}
library(gapminder)
```

```{r}
library(dplyr)
```


```{r}
gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007)+
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.4)
```
```{r}
ggplot(gapminder_2007)+
  aes(x=gdpPercap, y=lifeExp, color=pop)+
  geom_point(alpha=0.8)
```
```{r}
ggplot(gapminder_2007)+
  aes(x=gdpPercap, y=lifeExp, size=pop)+
  geom_point(alpha=0.5)
```
```{r}
ggplot(gapminder_2007)+
  geom_point(aes(x=gdpPercap, y=lifeExp, size=pop), alpha=0.5) +
  scale_size_area(max_size=10)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)
```
```{r}
ggplot(gapminder_1957)+
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)+
  geom_point(alpha=0.7)+
  scale_size_area(max_size=15)
```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957|year==2007)
```
```{r}
ggplot(gapminder_1957)+
  geom_point(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop),alpha=0.7)+
  scale_size_area(max_size=15)+
  facet_wrap(~year)
```
